---
title: "Handwritten Digits Network"
output:
  html_document:
    code_folding: none
    number_sections: yes
    toc: yes
    toc_float: yes
  html_notebook: 
    code_folding: none
    number_sections: yes
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: inline
runtime: shiny
---

<style>
button.hidecode + pre {display: none}
button.hideoutput + pre + pre {display: none}
</style>

# Used libraries and includes
```{r setup, cache = TRUE, message = FALSE, warning = FALSE}
library(keras)
library(ggplot2)
library(shiny)
```

# Session Info
<button class="hidecode hideoutput" onclick="$(this).next().toggle(); $(this).next().next().toggle();">Session Info</button>
```{r session_info, cache = TRUE}
sessionInfo()
(keras::backend())$backend()
```

# Loading and Visualizing Data
Loading data:
```{r loading_data, cache = TRUE}
image_size <- 400
data <- R.matlab::readMat('data/data.mat')
X <- data[['X']]
y <- data[['y']]
n <- nrow(X)
```

Randomly select 100 data points to display:
<button class="hidecode" onclick="$(this).next().toggle();">Display Data Array Function</button>
```{r display_data_array_function}
rotate <- function(x) t(apply(x, 2, rev))

displayDataArray <- function(sel) {
  n <- sqrt(nrow(sel))
  width <- sqrt(nrow(matrix(sel[1,])))
  height <- width
  full_img <- matrix(nrow = 0, ncol = width * n)
  for (i in 1:n) {
    row <- matrix(nrow = height, ncol = 0)
    for (j in 1:n) {
      img <- matrix(sel[n * (i - 1) + j,], nrow = height, ncol = width)
      row <- cbind(row, img)
    }
    full_img <- rbind(full_img, row)
  }
  full_img <- rotate(full_img)
  image(full_img, axes = FALSE, col = grey(seq(0, 1, length = 256))) 
}
```
```{r display_data}
displayDataArray(X[sample(nrow(X),size = 100),])
```

# Keras
## Already trained
### Loading Parameters
Load the weights into variables Theta1 and Theta2
```{r loading_weights, cache = TRUE}
weights <- R.matlab::readMat('data/weights.mat')
Theta1 <- weights[['Theta1']]
Theta2 <- weights[['Theta2']]
```

### Load Weight Functions
<button class="hidecode" onclick="$(this).next().toggle();">Predict Function</button>
```{r loading_weights_functions}
set_layer_weights <- function(model, index, theta) {
  current_weights <- model %>% get_layer(index = index) %>% get_weights()
  current_weights[[1]] = t(theta[,2:ncol(theta)])
  current_weights[[2]][] = c(theta[,1])[]
  model %>% get_layer(index = index) %>% set_weights(current_weights)
}
```

### Create Model And Load Weights
```{r create_model_with_predefined_weights}
model <- keras_model_sequential() 
model %>% 
  layer_dense(units = 25, activation = "sigmoid", input_shape = c(image_size)) %>%
  layer_dense(units = 10, activation = "sigmoid")

model %>% set_layer_weights(1, Theta1)
model %>% set_layer_weights(2, Theta2)

summary(model)

model %>% compile(
  loss = "categorical_crossentropy",
  optimizer = optimizer_adamax(),
  metrics = c("accuracy")
)
```

### Evuluate Model
```{r}
accuracy <- (model %>% evaluate(X, to_categorical(y - 1, num_classes = 10),verbose = 0))$acc
```
Training Set Accuracy: `r accuracy`

## Implement Predict By Hand

### Predict Function
<button class="hidecode" onclick="$(this).next().toggle();">Predict Function</button>
```{r predict_function}
sigmoid <- function(z) {
  return(1.0 / (1.0 + exp(-z)))
}

predict <- function(Theta1, Theta2, X) {
  m <- nrow(X)
  extendedX <- cbind(matrix(1, m, 1), X)
  z2 <- extendedX %*% t(Theta1)
  a2 <- sigmoid(z2)
  extendeda2 <- cbind(matrix(1, m, 1), a2)
  z3 <- extendeda2 %*% t(Theta2)
  h <- sigmoid(z3)
  
  return(matrix(apply(h, 1, which.max)))
}
```

### Evuluate Model
```{r}
accuracy <- mean(as.double(apply(predict(Theta1, Theta2, X),2,'==',y))) * 100
```
Training Set Accuracy: `r accuracy`

## My Model
### Dense Model
#### Prepare Data
```{r, cache = TRUE}
n_train <- round(0.8 * n)
shuffled <- sample(n);
x_all <- X[shuffled,]
y_all <- to_categorical(y[shuffled,] - 1, num_classes = 10)
x_train <- x_all[1:n_train,]
y_train <- y_all[1:n_train,]
x_test <- x_all[(n_train + 1):n,]
y_test <- y_all[(n_train + 1):n,]
```

#### Create model and compile
```{r}
model <- keras_model_sequential() 
model %>% 
  layer_dense(units = 512, activation = "sigmoid", input_shape = c(image_size)) %>%
  layer_dense(units = 10, activation = "softmax")

model %>% compile(
  loss = "categorical_crossentropy",
  optimizer = optimizer_nadam(),
  metrics = c("accuracy")
)

summary(model)
```

#### Training and Evaluation
```{r eval = 4, echo = 1}
history <- model %>% fit(
  x_train,
  y_train, 
  epochs = 30,
  batch_size = 32, 
  validation_data = list(x_test, y_test),
  shuffle = TRUE,
  #callbacks = callback_tensorboard("logs/run_a"),
  verbose = 2
)

plot(history)

model %>% save_model_hdf5("data/model.h5")
model <- load_model_hdf5("data/model.h5")
```

#### Evaluate
```{r}
accuracy <- (model %>% evaluate(x_test, y_test,verbose = 0))$acc
```
Training Set Accuracy: `r accuracy`

#### Generate prediction
```{r}
prediction_dense <- model %>% predict_classes(x_test)
```

### Convolution Model
#### Prepare Data
```{r, cache = TRUE}
n_train <- round(0.8 * n)
shuffled <- sample(n);
x_prepare <- array(X[shuffled,], dim=c(n, 20, 20))
x_all <- array(rep(1, length(x_prepare)), dim=c(n, 20, 20, 3))
for (i in 1:3) {
  x_all[,,,i] <- x_prepare
}
y_all <- to_categorical(y[shuffled,] - 1, num_classes = 10)
x_train <- x_all[1:n_train,,,]
y_train <- y_all[1:n_train,]
x_test <- x_all[(n_train + 1):n,,,]
y_test <- y_all[(n_train + 1):n,]
dim(x_test)
dim(y_test)
```

#### Create model and compile
layer_max_pooling_2d - Reduces the dimentions and the computational complexity
layer_dropout - Dropout layer to avoid overfitting
layer_flatten - Flatten the input 
```{r}
model <- keras_model_sequential() 
model %>% 
  layer_conv_2d(filter = 32, kernel_size = c(3, 3), activation = "relu", input_shape = c(20, 20, 3)) %>% #, padding = "same"
  layer_conv_2d(filter = 64, kernel_size = c(3, 3), activation = "relu")  %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_dropout(0.25) %>%
  layer_flatten() %>%
  layer_dense(128, activation = "relu") %>%
  layer_dropout(0.5) %>%
  layer_dense(10, activation = "softmax")

model %>% compile(
  loss = "categorical_crossentropy",
  optimizer = optimizer_adadelta(), #optimizer_adam(lr = 0.0001 , decay = 1e-6), #lr - learning rate, decay - learning rate decay over each update
  metrics = c("accuracy")
)

summary(model)
```

#### Training and Evaluation
```{r eval = 4, echo = 1}
history <- model %>% fit(
  x_train,
  y_train, 
  epochs = 30,
  batch_size = 128, 
  validation_data = list(x_test, y_test),
  shuffle = TRUE
)

plot(history)

model %>% save_model_hdf5("data/CNN.h5")
model <- load_model_hdf5("data/CNN.h5")
```

#### Generating images
```{r}
gen_images <- image_data_generator(
  #featurewise_center = TRUE,
  #featurewise_std_normalization = TRUE,
  rotation_range = 15,
  width_shift_range = 0.15,
  height_shift_range = 0.15,
  zoom_range = 0.15
)
#Fit image data generator internal statistics to some sample data
gen_images %>% fit_image_data_generator(x_train)
#Generates batches of augmented/normalized data from image data and #labels to visually see the generated images by the Model
batches <- flow_images_from_data(
  x_all, 
  y_all, 
  gen_images,
  batch_size = 32
  #save_to_dir="~/HandwrittenDigitsNetwork/generated_images/"
)
model %>% fit_generator(
     batches,
     steps_per_epoch = as.integer(4000/32),
     epochs = 30,
     validation_data = list(x_test, y_test)
)

model %>% save_model_hdf5("data/CNN.h5")
```

#### Evaluate
```{r}
accuracy <- (model %>% evaluate(x_test, y_test,verbose = 0))$acc
```
Training Set Accuracy: `r accuracy`

#### Generate prediction
```{r}
prediction_conv <- model %>% predict_classes(x_test)
```

# Test

## By Hand
<button class="hidecode" onclick="$(this).next().toggle();">Display Data Function</button>
```{r display_data_function}
displayData <- function(sel) {
  width <- sqrt(ncol(sel))
  mat <- matrix(sel, nrow = width, ncol = width)
  mat <- rotate(mat)
  image(mat, axes = FALSE, col = grey(seq(0, 1, length = 256))) 
}
```

```{r test}
displayRandomImageAndGetPrediction <- function(Theta1, Theta2, X) {
    # Select random image
    example <- t(matrix(X[sample(nrow(X), 1),]));
    
    # Display
    displayData(example);
    
    # Prediction
    return(c(predict(Theta1, Theta2, example)))
}
```

<button class="hidecode" onclick="$(this).next().toggle();">UI Functions</button>
```{r}
renderPlot({
  input$updateButton
  
  prediction <- displayRandomImageAndGetPrediction(Theta1, Theta2, X)
  output$prediction <- renderText({ 
    paste('Neural Network Prediction: ', prediction, '\n')
  })
}, width = 200, height = 200)
actionButton('updateButton', 'Update')
textOutput('prediction')
```

## My Dense Model
```{r}
displayRandomImageAndGetPredictionKeras <- function(x, y) {
    # Select random image
    i <- sample(nrow(x), 1)
    example <- t(matrix(x[i,]));
    
    # Display
    displayData(example);
    
    # Prediction
    return(y[i])
}
```

<button class="hidecode" onclick="$(this).next().toggle();">UI Functions</button>
```{r}
renderPlot({
  input$updateButton2
  
  prediction <- displayRandomImageAndGetPredictionKeras(x_test, prediction_dense)
  output$prediction2 <- renderText({ 
    paste('Neural Network Prediction: ', prediction, '\n')
  })
}, width = 200, height = 200)
actionButton('updateButton2', 'Update')
textOutput('prediction2')
```

# License
MIT License

Copyright (c) 2017 Konstantin Stukov

Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files (the “Software”), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED “AS IS”, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.